### 2021-03-18 Paper.tex
template

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Software Product Lines and Variability\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits.

\section{Ease of Use}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2021-03-18 Paper.tex
authors, ignore pdf

@@
 
-\title{Software Product Lines and Variability\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
-}
+\title{Software Product Lines and Variability}
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
+\author{\IEEEauthorblockN{Niklas Hauber}
+\IEEEauthorblockA{\textit{Johannes Kepler University} \\
+Linz, Austria \\
+niklas.hauber@solentia.at}
 \and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
+\IEEEauthorblockN{Jonas Reichhardt}
+\IEEEauthorblockA{\textit{Johannes Kepler University} \\
 City, Country \\
@@
 }
-
 \maketitle

### 2021-03-18 Paper.tex
added mail address
added abstract and sections

@@
 \documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
 \usepackage{cite}
@@
 
-\title{Software Product Lines and Variability}
+\title{Managing variability in software systems}
 
@@
 \IEEEauthorblockA{\textit{Johannes Kepler University} \\
-City, Country \\
-email address or ORCID}
+Linz, Austria\\
+j.reichhardt@tig.at}
 }
@@
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+With growing complexity of software systems, variability of said has become an increasing problem and managing it is thus a crucial part of the present-day software engineering process.
+It is particularly vital if either many product versions make use of an existing architecture over a long period of time or if a variety of software systems share a common code base.
+The focus of this paper is to provide an overview of how variation in software systems is caused, how it influences the development process and how it can be dealt with.
 \end{abstract}
 
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
+\section{Introduction}
+Here we are going to introduce the reader into the broad environment
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
+\section{Background and related work}
+In this section the background research which we use to present an overview of the topic.
 
-\section{Ease of Use}
+\section{Origin of variability}
+We introduce why there is variability in the modern software engineering process. So that the reader knows how the problems arrise to have a baseline for presenting solutions for said problems.
 
-\subsection{Maintaining the Integrity of the Specifications}
+\section{Variability modeling}
+The question we want to answer in this section how one can model/describe the changes the software architecture is exposed to and how the dependencies and flexibility can be put into a defined state.
 
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
+\section{Variability patterns}
+Sums up the typical patterns how variability can be introduced.
 
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
+\section{Discussion/Challenges}
+Discuss why the variability of systems is hard to tackle and why it is not always successful.
 
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-\begin{figure}[htbp]
-\centerline{\includegraphics{fig1.png}}
-\caption{Example of a figure caption.}
-\label{fig}
-\end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
+\section{Conclussion}
+Make final remarks about the discussed elements.
 
 \begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
+\bibitem{b1} Apel S., Batory D., Kästner C., Saake G. (2013) Software Product Lines. In: Feature-Oriented Software Product Lines. Springer, Berlin, Heidelberg. ISBN 978-3-642-37520-0
+\bibitem{b2} J.
+van Gurp, J.
+Bosch and M.
+Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture, Amsterdam, Netherlands, 2001, pp.
+45-54, doi: 10.1109/WICSA.2001.948406.
+\bibitem{b3} Bachmann F., Bass L. (2001). Managing Variability in Software Architectures. Software Engineering Institute, Carnegie Mellon University.
+\bibitem{b4} Galster M., Weyns D., Tofan  D., Bartosz M., Avgeriou P.
+(2014).
+Variability in Software Systems—A Systematic Literature Review.
+Software Engineering, IEEE Transactions on.
+40.
+282-306.
+10.1109/TSE.2013.56.
 \end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
 

### 2021-04-22 Paper.tex
added variability planning
divide section workload

@@
 Here we are going to introduce the reader into the broad environment  
+(1/2 page) JR
 
@@
 In this section the background research which we use to present an overview of the topic.
+(1/2 page) NH
 
@@
 We introduce why there is variability in the modern software engineering process. So that the reader knows how the problems arrise to have a baseline for presenting solutions for said problems.
+(1/2 page) NH
 
@@
 The question we want to answer in this section how one can model/describe the changes the software architecture is exposed to and how the dependencies and flexibility can be put into a defined state. 
+(1 page) JR
 
@@
 Sums up the typical patterns how variability can be introduced.
+(1/2 page, maybe merge with section above) NH
+
+\section{???Variability planning???}
+Gives an overview of how someone can plan for varability and how to manage it.
 
@@
 Discuss why the variability of systems is hard to tackle and why it is not always successful.
+(1/4 page) JR
 
@@
 Make final remarks about the discussed elements.
+(1/4 page) NH
 
 \begin{thebibliography}{00}
-\bibitem{b1} Apel S., Batory D., Kästner C., Saake G. (2013) Software Product Lines. In: Feature-Oriented Software Product Lines. Springer, Berlin, Heidelberg. ISBN 978-3-642-37520-0
+\bibitem{b1} S. Apel, D. Batory, C. Kästner and G. Saake, "Software Product Lines. In: Feature-Oriented Software Product Lines", Berlin Heidelberg, Springer (2013)
 \bibitem{b2} J.
@@
 Bosch and M.
-Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture, Amsterdam, Netherlands, 2001, pp.
-45-54, doi: 10.1109/WICSA.2001.948406.
-\bibitem{b3} Bachmann F., Bass L. (2001). Managing Variability in Software Architectures. Software Engineering Institute, Carnegie Mellon University.
-\bibitem{b4} Galster M., Weyns D., Tofan  D., Bartosz M., Avgeriou P.
-(2014).
-Variability in Software Systems—A Systematic Literature Review.
-Software Engineering, IEEE Transactions on.
-40.
-282-306.
-10.1109/TSE.2013.56.
+Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture", Amsterdam, Netherlands, 2001, pp.
+45-54
+\bibitem{b3} F. Bachmann, L. Bass, "Managing Variability in Software Architectures", Carnegie Mellon University (2001)
+\bibitem{b4} M. Galster, D. Weyns, D. Tofan, M. Bartosz, P. Avgeriou, "Variability in Software Systems—A Systematic Literature Review", IEEE Transactions on Software Engineering (2014)
+\bibitem{b5} T.
+Berger, J.-P.
+Steghöfer, T.
+Ziadi, J.
+Robin, and J.
+Martinez, "The state of adoption and the challenges of systematic variability management in industry", Empirical Software Engineering, vol.
+25, pp.
+1755-1797, 2020
 \end{thebibliography}

### 2021-04-24 Paper.tex
add Introduction v1

@@
 \section{Introduction}
-Here we are going to introduce the reader into the broad environment  
+Building software systems which can adopt to changes from outside as well as within said system represents a vital goal to build software systems that last.
+When these goals are achieved the effort and therefore costs are reduced for maintenance, but also for further improvements and extension.
+\\\\
+Therefore, a lot of companies design their software such as numerous variants exists, which allows customization to meet the requirements which are laid out by customers and stakeholders.
+One of the most used practices to achieve a variant-rich software system is software product line engineering (SPLE) which uses concept, processes, modeling techniques for features and function, design patters and respective tools to support the process.[5]\\\\
+If one wants to design a system which is prepared for change, the focus should lie on developing components which have loose coupling and strong cohesion.
+So a module solves one simple task and is as independent as it can be.
+This leads to the least amount of unwanted side effects in other modules if one module is confronted with unpredicted variability.\\\\
+We will try to provide a basic introduction how variability is managed in software systems by collection information from other studies which describe the key aspects of this vast field of research.
+In the first section we are providing some base knowledge acquired previous studies.
+After that we will lay out how one can identify the various sources of variability in software systems.
+After that we will bring up techniques and processes which have the purpose of modeling the variation in a software system.
+Last but not least we will discuss some of the big tasks which have to be addressed and boil down the described aspects in this paper to a conclussion.
 (1/2 page) JR

### 2021-04-25 Paper.tex
background & related work

@@
 \begin{abstract}
-With growing complexity of software systems, variability of said has become an increasing problem and managing it is thus a crucial part of the present-day software engineering process.
+With the growing complexity of software systems, variability of said has become an increasing problem and managing it is thus a crucial part of the present-day software engineering process.
 It is particularly vital if either many product versions make use of an existing architecture over a long period of time or if a variety of software systems share a common code base.
@@
 \section{Background and related work}
-In this section the background research which we use to present an overview of the topic.
+As variability has always been a problem in software systems, a lot of research and developement have been put into this topic, in order to advance the subject of managing variability.\\
+
+Managing variability is certainly not a trivial task, as requirements differ vastly when looking at different industries.
+Therefore several strategies exist for engineering variant-rich systems, each with its own advantages and disadvantages.
+\cite{b5}
+
+The main direction existing research has progressed is how to model variability \cite{b4}, by attempting to formulate principles of building and maintaining said models.
+Often by analyzing various existing papers on the topic.
+Two approaches of how to model variability have gained the most importance: feature modeling and decision modeling.
+\\
+
+The origin of feature modeling is the feature domain analysis method (FODA), which tries to outline common and variable parts of a system.
+A lot of work as since been done on this, extending the original notation.\cite{b6}\\
+
+Decision modeling has its earliest documented approach in the Synthesis method.
+\cite{b6} This defines it as \textit{a set of decisions that are adequate to distinguish among the members of an application engineering product family and to guide adaptation of application
+	engineering work products}. \cite{b7}\\
+
 (1/2 page) NH
@@
 1755-1797, 2020
+\bibitem{b6}
+K. Czarnecki, P. Grünbacher, R. Rabiser, K. Schmid., and A.
+Wąsowski, “Cool features and tough decisions: a comparison of
+variability modeling approaches,” Proc. of the \textit{Sixth international
+workshop on variability modeling of software-intensive systems
+(VaMoS '12)}, ACM, Jan 2012, pp. 173-182.
+\bibitem{b7}
+Software Productivity Consortium Services Corporation,
+Technical Report SPC-92019-CMC. \textit{Reuse-Driven Software
+Processes Guidebook}, Version 02.00.03, 1993.
+
 \end{thebibliography}

### 2021-04-28 Paper.tex
start with variability modeling

@@
 Therefore, a lot of companies design their software such as numerous variants exists, which allows customization to meet the requirements which are laid out by customers and stakeholders.
-One of the most used practices to achieve a variant-rich software system is software product line engineering (SPLE) which uses concept, processes, modeling techniques for features and function, design patters and respective tools to support the process.[5]\\\\
+One of the most used practices to achieve a variant-rich software system is software product line engineering (SPLE) which uses concept, processes, modeling techniques for features and function, design patters and respective tools to support the process.\cite{b5}\\\\
 If one wants to design a system which is prepared for change, the focus should lie on developing components which have loose coupling and strong cohesion.
@@
 Last but not least we will discuss some of the big tasks which have to be addressed and boil down the described aspects in this paper to a conclussion.
-(1/2 page) JR
 
-\section{Background and related work}
+\section{Background and related work (1/2 page) NH}
 As variability has always been a problem in software systems, a lot of research and developement have been put into this topic, in order to advance the subject of managing variability.\\
@@
 
-(1/2 page) NH
+\section{Origin of variability (1/2 page) NH}
+We introduce why there is variability in the modern software engineering process. So that the reader knows how the problems arrise to have a baseline for presenting solutions for said problems.
 
-\section{Origin of variability}
-We introduce why there is variability in the modern software engineering process. So that the reader knows how the problems arrise to have a baseline for presenting solutions for said problems.
-(1/2 page) NH
-
-\section{Variability modeling}
-The question we want to answer in this section how one can model/describe the changes the software architecture is exposed to and how the dependencies and flexibility can be put into a defined state. 
-(1 page) JR
+\section{Variability modeling (1 page) JR}
+As mentioned there are two main approaches for managing variability.
+For each of which we will discuss the advantages and disantages and create a broad overview over the different modeling approaches.\\\\
+\textbf{Feature modeling}\\
+Features in the context of feature modeling are behaviours, characetristics or functionalities which are observable by the user of the software system.
+They can be used to define commonalities and differences between systems and their variants, as well as determine mandatory, optional or alternative aspects of them\cite{b8}.
+A broadly used approach of displaying features and their relations are hierarchically structured diagrams which are called features diagrams.\\\\
+A big advantage of feature models, especially in the shape of feature diagrams, is a possible communication enhancement between stakeholders.
+This can be achieved by providing a overview of all functionalities which are offered to the user by the system.
+Due to the nature of the model regarding structure and due to the fact that it displays if and which planned features are already provided by another system or system family, this may improve the cost and time estimation for developing the software system.
+\\\\
+\textbf{Decision modeling}\\
+asdfasdf
 
@@
 K. Czarnecki, P. Grünbacher, R. Rabiser, K. Schmid., and A.
-Wąsowski, “Cool features and tough decisions: a comparison of
+Wasowski, “Cool features and tough decisions: a comparison of
 variability modeling approaches,” Proc. of the \textit{Sixth international
@@
 Processes Guidebook}, Version 02.00.03, 1993.
+\bibitem{b8} K.
+C.
+Kang, S.
+Cohen, J.
+Hess, W.
+Novak, and S.
+Peterson,	 "Feature oriented domain analysis (FODA) feasibility study", Technical Report CMU/SEI-90-TR-21, Software Engineering Institute, Carnegie Mellon University, Pittburgh, PA, USA, 1990
 

### 2021-04-29 Paper.tex
work on feature modeling

@@
 As mentioned there are two main approaches for managing variability.
-For each of which we will discuss the advantages and disantages and create a broad overview over the different modeling approaches.\\\\
+For each of which we will discuss the advantages and disadvantages and create a broad overview over the different modeling approaches.\\\\
 \textbf{Feature modeling}\\
-Features in the context of feature modeling are behaviours, characetristics or functionalities which are observable by the user of the software system.
+Features in the context of feature modeling are behaviours, characteristics or functionalities which are observable by the user of the software system.
 They can be used to define commonalities and differences between systems and their variants, as well as determine mandatory, optional or alternative aspects of them\cite{b8}.
-A broadly used approach of displaying features and their relations are hierarchically structured diagrams which are called features diagrams.\\\\
+A broadly used approach of displaying features and their relations is hierarchically structured diagram which is called a features diagrams.\\\\
 A big advantage of feature models, especially in the shape of feature diagrams, is a possible communication enhancement between stakeholders.
@@
 \\\\
+The essence of feature modeling is the splitting of features into distinct objects which have a certain relation to each other.
+There are several types of features mentioned in \cite{b8}, named \textit{mandatory, optional and alternative features}.
+Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.\cite{b9}\\\\
+If a feature is mandatory it has to be selected if its parent feature is selected, the type of the parent feature does not matter.
+A optional feature can be selected freely but there may be constraints from composition rules (e.g.
+a feature can only be choosen as part of a feature sets).
+Alternative features are can be pictured as a choice box, so one can choose \textit{one} item from a feature set.\\\\
+This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features.
+\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.
+
 \textbf{Decision modeling}\\
@@
 Peterson,	 "Feature oriented domain analysis (FODA) feasibility study", Technical Report CMU/SEI-90-TR-21, Software Engineering Institute, Carnegie Mellon University, Pittburgh, PA, USA, 1990
-
+\bibitem{b9} A.
+Metzger, and K.
+Pohl, "Software product line engineering and variability management: Achievements and challenges," \textit{Proc.
+of the Future of Software Engineering (FOSE 2014)}, ACM, 2014, pp.
+70-84, doi:10.1145/2593882.2593888.
+\bibitem{b10} M.
+Riebisch, K.
+Böllert, D.
+Streitferdt, and I.
+Philippow, "Extending Feature Diagrams with UML Multiplicities", \textit{6th World Conference on Integrated Design \& Process Technology (IDPT2002)}, vol.
+23, 2002, pp.
+1-7.
 \end{thebibliography}

### 2021-04-30 Paper.tex
added graph

@@
 \usepackage{xcolor}
+\usepackage{svg}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features.
-\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.
+\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.\\\\
+In Fig.
+1, one can see a small feature diagram of a car, it consists of one root node, that describes the domain we are modeling, in this case the car.
+The previous elaborated mandatory features are the engine and seats.
+Furthermore there is the option to install a cruise control and the option for a turbocharger these features are completely independet and optional.
+Last but not least one can choose between normal leather seats or alcantara ones but not both.
+Which this simplified model of a car we have all three types of features modeled
 

### 2021-05-02 Paper.tex
origins of variability

@@
 \cite{b6} This defines it as \textit{a set of decisions that are adequate to distinguish among the members of an application engineering product family and to guide adaptation of application
-	engineering work products}. \cite{b7}\\
+	engineering work products}. [7 p. 104]\\
+
+All in all there exists various literature on the different scopes variability management encompasses and we try to capture the main points of several of them in this paper.
 
 \section{Origin of variability (1/2 page) NH}
-We introduce why there is variability in the modern software engineering process. So that the reader knows how the problems arrise to have a baseline for presenting solutions for said problems.
+Variability is the ability of software to be changed for specific context.
+That encompasses extension, customization and reconfiguration.
+It thus defines the parts of the software that remain variable and are not fully specified yet.
+\cite{b4}\\\\
+There are two main cases where one needs to capture different alternatives within an architecture:\\
+\begin{enumerate}
+	\item At the time of design there may exist various alternatives and if a decision cannot be made yet because of lack of information, it will be deferred.
+Therefore one must design the system to be able to handle all viable alternatives.\\
+	\item There already exists an product line which contains several different alternatives for given parts of software, which can be chosen depending on the needs of each individual customer.\\
+\end{enumerate}
+While these situations differ greatly in reality, in the context of variability management they are really the same.
+That is, before a decision is made on which alternative is going to be chosen, one does not know the outcome and needs to be able to handle all possible ones.\cite{b3}\\
 
+There are also several reasons why such variations have to exist in the first place \cite{b3}:
+\begin{enumerate}
+	\item \textit{Variation in function.} Not every product supports the same functions.
+Take as example a modern smartphone, some might support fingerprint authentication, some might support facial recognition authentication and some might support none of those.\\
+	\item \textit{Variation in data.} Products might vary in the definitions of data structures.
+For example a software managing planned repairs of industrial machines might need store additional information for some companies like an SAP Id in case these planned repairs are also concurrently managed in SAP, while some other companies might not need that information.\\
+	\item \textit{Variation in technology.} The technology different products run on may vary.
+Some software systems for example need to run on Linux while others need to support Windows.
+Or for some a simple MySQL database might suffice while others need the SQL Server Enterprise edition because the amount of data might vary significantly.\\
+	\item \textit{Variation in environment.} The environment software exists in might vary. Meaning also the way in which it interacts with its environment. \\
+\end{enumerate}
+These, at least, are the four most common origins of variation. Of course some more do exist, but these are beyond the scope of this document.\\
 \section{Variability modeling (1 page) JR}
@@
 \section{Variability patterns}
-Sums up the typical patterns how variability can be introduced.
-(1/2 page, maybe merge with section above) NH
+Variability can be introduced into a system via different mechanisms.
+These mechanisms strive to bind a given system to one out of many alternatives at different binding times.
+Binding times describe when exactly these mechanisms stop deffering the decision and make the choice for any of the variations, as illustrated in Figure 1.\cite{b2}
 
+\begin{figure}[h!]
+	\includegraphics[width=\columnwidth]{binding_times}
+	\caption{Binding Times \cite{b2}}
+\end{figure}
+
+Further explanation of these binding times can be found in the subsequent paragraphs:\\
+\begin{itemize}
+	\item \textit{Product Architecture Derivation.}
+\end{itemize}
 \section{???Variability planning???}

### 2021-05-02 Paper.tex
variability patterns conclusion

@@
 \begin{itemize}
-	\item \textit{Product Architecture Derivation.}
+	\item \textit{Product architecture derivation.} The architecture of the product already binds to a variation. This is the lowest binding time and will influence all subsequent steps.\\
+	\item \textit{Compilation.} When the source code is compiled there exist several options to bind to a variation by adjusting constants, macros or compiler directives.\\
+	\item \textit{Linking.} This option does not exist for any technology, but for many programming languages the linking process can be manually changed. For example by switching the DLL of a module.\\
+	\item \textit{Startup} A simple option to bind variants to the system is by configuration files, that influence the features a software system exhibits.\\
+	\item \textit{Adaptation to runtime environment} This typically refers to plugins of a software, which can add variation to a system and can either be provided by the original developer or even third party vendors.
+\\
 \end{itemize}
-\section{???Variability planning???}
-Gives an overview of how someone can plan for varability and how to manage it.
-
+Any mechanism to bind to none, one or many different variants tend to fall into different categories presented by \cite{b2} and \cite{b3}, and also have parallels to the feature types presented in the previous section:\\
+\begin{itemize}
+	\item \textit{Variant Entity.} There exist several options but only one may be bound to at any given time.\\
+	\item \textit{Optional Entity.} As the name suggests, an optional entity may or may not be included in the system.\\
+	\item \textit{Multiple Coexisting Entities.} Similar to the variant entity but with the exception that multiple variants can be used.\\
+\end{itemize}
+These categories tend to be the same for the different binding times, but of course vary greatly in their implementation.
+While an optional entity during compilation may only be influenced by a compiler flag, it might be a whole plugin from a third party vendor when adapting a given runtime environment.
+An in depth analysis of these mechanisms can be found in \cite{b2}.
 \section{Discussion/Challenges}
@@
 
-\section{Conclussion}
-Make final remarks about the discussed elements.
-(1/4 page) NH
+\section{Conclussion (1/4 page) NH}
+In this document we have given a short introduction to the origin of variability, how to model it and what patterns exist as well as look over related work to these topics. \\
+
+This information is vital to being able to manage variability, it does however, only glance over the different subjects and only provides a theoretical basis.
+Practical use also depends on supporting software, which widely exists especially for the task of variability modelling, as well as practical knowledge in the field, which also aids in gaining an in depth understanding of managing variability.\\
+
+For further information we refer the reader to the references, especially \cite{b2}, which gives a more in depth overview, and \cite{b6}, which compares different approaches to variability modelling.
+\newpage
 

### 2021-05-02 Paper.tex
add discussion

--- 2021-04-30 Paper.tex
@@
 Alternative features are can be pictured as a choice box, so one can choose \textit{one} item from a feature set.\\\\
-This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features.
-\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.\\\\
+This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features.\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.\\\\
 In Fig.
@@
 Last but not least one can choose between normal leather seats or alcantara ones but not both.
-Which this simplified model of a car we have all three types of features modeled
-
+Which this simplified model of a car we have all three types of features modeled.\\\\
 \textbf{Decision modeling}\\
-asdfasdf
+In decision modeling the variability is described with decision instead of features.
+A variable feature in feature modeling is equal to a decision and vice versa.
+Contrary to feature modeling decision modeling is unable to describe the commonalities of systems.\\\\
+Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependicies.
+In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy.
+\cite{b6}\\\\
+Same as feature modeling, decision modeling has been expanded over time to enable the application coverage of DM be almost the same as FM.\cite{b6}\\\\
 
@@
 
-\section{Discussion/Challenges}
-Discuss why the variability of systems is hard to tackle and why it is not always successful.
-(1/4 page) JR
+\section{Discussion/Challenges (1/4 page) JR}
+The research field of variability had a increasing trend of published paper per year from 2000 until 2008.
+After that it stabilized between 25-30 new research paper per year from 2008 until 2011.
+One explanation of this rise in interest in the scientific community could be that the usage of variability was becoming more broadly in dynamic systems.
+\cite{b4}\\\\
+Another aspect is that feature modeling is in its core a simple concept.
+Even though its impact in the industry is pretty big considering it is the most popular concept for variability modeling in software systesm until today.\cite{b9} One can say that because it is such a simple concept and due to the big amount of research in this field we understand the subject pretty well by now.\\\\
+The currently existing methods also have their limitations.
+It is difficult to guarantee correctness and consistency, authors have to urge for verification and consistency checking.
+Performance overheads, runtime upgradability and the lack of support regarding security needs are some areas where the described methods are limited.
+Sometimes the handling of variability can create poor user-friendliness and therefore decrease usebility of a system.
+\cite{b4}\\\\
+There are ways to decrease the negative impact of these aspects e.g.
+correctness and consistency can be checked by various methods such as modeling guidelines customized for specific needs, together with a reviewing process where changes to the model are checked by another person to ensure the quality of the metioned parameters.
 

### 2021-05-02 Paper.tex
merge conflict

--- 2021-05-02 Paper.tex
@@
 An in depth analysis of these mechanisms can be found in \cite{b2}.
-\section{Discussion/Challenges}
-Discuss why the variability of systems is hard to tackle and why it is not always successful.
-(1/4 page) JR
+
+\section{Discussion/Challenges (1/4 page) JR}
+The research field of variability had a increasing trend of published paper per year from 2000 until 2008.
+After that it stabilized between 25-30 new research paper per year from 2008 until 2011.
+One explanation of this rise in interest in the scientific community could be that the usage of variability was becoming more broadly in dynamic systems.
+\cite{b4}\\\\
+Another aspect is that feature modeling is in its core a simple concept.
+Even though its impact in the industry is pretty big considering it is the most popular concept for variability modeling in software systesm until today.\cite{b9} One can say that because it is such a simple concept and due to the big amount of research in this field we understand the subject pretty well by now.\\\\
+The currently existing methods also have their limitations.
+It is difficult to guarantee correctness and consistency, authors have to urge for verification and consistency checking.
+Performance overheads, runtime upgradability and the lack of support regarding security needs are some areas where the described methods are limited.
+Sometimes the handling of variability can create poor user-friendliness and therefore decrease usebility of a system.
+\cite{b4}\\\\
+There are ways to decrease the negative impact of these aspects e.g.
+correctness and consistency can be checked by various methods such as modeling guidelines customized for specific needs, together with a reviewing process where changes to the model are checked by another person to ensure the quality of the metioned parameters.
 

### 2021-05-02 Paper.tex
added itemize

--- 2021-05-02 Paper.tex
@@
 
-\section{Background and related work (1/2 page) NH}
+\section{Background and related work}
 As variability has always been a problem in software systems, a lot of research and developement have been put into this topic, in order to advance the subject of managing variability.\\
@@
 
-\section{Origin of variability (1/2 page) NH}
+\section{Origin of variability}
 Variability is the ability of software to be changed for specific context.
@@
 These, at least, are the four most common origins of variation. Of course some more do exist, but these are beyond the scope of this document.\\
-\section{Variability modeling (1 page) JR}
+\section{Variability modeling}
 As mentioned there are two main approaches for managing variability.
@@
 They can be used to define commonalities and differences between systems and their variants, as well as determine mandatory, optional or alternative aspects of them\cite{b8}.
-A broadly used approach of displaying features and their relations is hierarchically structured diagram which is called a features diagrams.\\\\
+A broadly used approach of displaying features and their relations is hierarchically structured diagram which is called a features diagrams.\\
+
 A big advantage of feature models, especially in the shape of feature diagrams, is a possible communication enhancement between stakeholders.
@@
 Due to the nature of the model regarding structure and due to the fact that it displays if and which planned features are already provided by another system or system family, this may improve the cost and time estimation for developing the software system.
-\\\\
+\\
+
 The essence of feature modeling is the splitting of features into distinct objects which have a certain relation to each other.
 There are several types of features mentioned in \cite{b8}, named \textit{mandatory, optional and alternative features}.
-Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.\cite{b9}\\\\
+Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.\cite{b9}\\
+
 If a feature is mandatory it has to be selected if its parent feature is selected, the type of the parent feature does not matter.
@@
 a feature can only be choosen as part of a feature sets).
-Alternative features are can be pictured as a choice box, so one can choose \textit{one} item from a feature set.\\\\
-This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features.
-\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.\\\\
+Alternative features are can be pictured as a choice box, so one can choose \textit{one} item from a feature set.\\
+
+This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features.\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.
+
+\begin{figure}[h!]
+	\includegraphics[width=\columnwidth]{feature_model}
+	\caption{Simplified feature model of a car}
+\end{figure}
 In Fig.
@@
 Last but not least one can choose between normal leather seats or alcantara ones but not both.
-Which this simplified model of a car we have all three types of features modeled
-
+Which this simplified model of a car we have all three types of features modeled.\\\\
 \textbf{Decision modeling}\\
-asdfasdf
+In decision modeling the variability is described with decision instead of features.
+A variable feature in feature modeling is equal to a decision and vice versa.
+Contrary to feature modeling decision modeling is unable to describe the commonalities of systems.\\\\
+Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependicies.
+In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy.
+\cite{b6}\\\\
+Same as feature modeling, decision modeling has been expanded over time to enable the application coverage of DM be almost the same as FM.\cite{b6}\\
 
@@
 These mechanisms strive to bind a given system to one out of many alternatives at different binding times.
-Binding times describe when exactly these mechanisms stop deffering the decision and make the choice for any of the variations, as illustrated in Figure 1.\cite{b2}
+Binding times describe when exactly these mechanisms stop deffering the decision and make the choice for any of the variations, as illustrated in Figure 2.\cite{b2}
 
@@
 An in depth analysis of these mechanisms can be found in \cite{b2}.
+
 \section{Discussion/Challenges}
-Discuss why the variability of systems is hard to tackle and why it is not always successful.
-(1/4 page) JR
+The research field of variability had a increasing trend of published paper per year from 2000 until 2008.
+After that it stabilized between 25-30 new research paper per year from 2008 until 2011.
+One explanation of this rise in interest in the scientific community could be that the usage of variability was becoming more broadly in dynamic systems.
+\cite{b4}\\
 
-\section{Conclussion (1/4 page) NH}
+Another aspect is that feature modeling is in its core a simple concept.
+Even though its impact in the industry is pretty big considering it is the most popular concept for variability modeling in software system until today.\cite{b9} One can say that because it is such a simple concept and due to the big amount of research in this field we understand the subject pretty well by now.\\
+
+As mentioned in \cite{b4} the currently existing methods also have their limitations. 
+\begin{itemize}
+\item \textit{Guaranteed correctness and consistency}. Authors have to urge for verification and consistency checking.\\
+\item \textit{Performance overheads} are difficult to model.\\
+\item \textit{Runtime upgradability} and the process of switching features is sometimes necessary.\\
+\item \textit{Lack of support for security needs} can be a factor whichs limits the use of feature modeling.\\
+\item \textit{Poor user-friendliness} can be introduced by poor handling of variability, which decreases the usebility of a system. As an example if you \\
+\end{itemize}
+
+There are ways to decrease the negative impact of these aspects e.g.
+correctness and consistency can be checked by various methods such as modeling guidelines customized for specific needs, together with a reviewing process where changes to the model are checked by another person to ensure the quality of the metioned parameters.\\
+
+It should also be noted most of the studies in the area of FM are including some sort of tool to support the process. Only a few don't mention any form of software support.\\
+
+\section{Conclussion}
 In this document we have given a short introduction to the origin of variability, how to model it and what patterns exist as well as look over related work to these topics. \\

### 2021-05-02 Paper.tex
proof reading

@@
 When these goals are achieved the effort and therefore costs are reduced for maintenance, but also for further improvements and extension.
-\\\\
+\\
+
 Therefore, a lot of companies design their software such as numerous variants exists, which allows customization to meet the requirements which are laid out by customers and stakeholders.
-One of the most used practices to achieve a variant-rich software system is software product line engineering (SPLE) which uses concept, processes, modeling techniques for features and function, design patters and respective tools to support the process.\cite{b5}\\\\
+One of the most used practices to achieve a variant-rich software system is software product line engineering (SPLE) which uses concept, processes, modeling techniques for features, functions, design patters and respective tools to support the process.\cite{b5}\\
+
 If one wants to design a system which is prepared for change, the focus should lie on developing components which have loose coupling and strong cohesion.
-So a module solves one simple task and is as independent as it can be.
-This leads to the least amount of unwanted side effects in other modules if one module is confronted with unpredicted variability.\\\\
-We will try to provide a basic introduction how variability is managed in software systems by collection information from other studies which describe the key aspects of this vast field of research.
+Thus a module solves one simple task and is as independent as it can be.
+This leads to the least amount of unwanted side effects in other modules if one module is confronted with unpredicted variability.\\
+
+We will try to provide a basic introduction of how variability is managed in software systems by collecting information from other studies which describe the key aspects of this vast field of research.
 In the first section we are providing some base knowledge acquired previous studies.
 After that we will lay out how one can identify the various sources of variability in software systems.
-After that we will bring up techniques and processes which have the purpose of modeling the variation in a software system.
-Last but not least we will discuss some of the big tasks which have to be addressed and boil down the described aspects in this paper to a conclussion.
+Then we will bring up techniques and processes which have the purpose of modeling the variation in a software system.
+Last but not least we will discuss some of the big tasks which have to be addressed and conclude the topic.
 
@@
 
-Decision modeling has its earliest documented approach in the Synthesis method.
+Decision modeling has its earliest documented approach in the synthesis method.
 \cite{b6} This defines it as \textit{a set of decisions that are adequate to distinguish among the members of an application engineering product family and to guide adaptation of application
@@
 Therefore one must design the system to be able to handle all viable alternatives.\\
-	\item There already exists an product line which contains several different alternatives for given parts of software, which can be chosen depending on the needs of each individual customer.\\
+	\item There already exists a product line which contains several different alternatives for given parts of the software, which can be chosen depending on the needs of each individual customer.\\
 \end{enumerate}
@@
 	\item \textit{Variation in function.} Not every product supports the same functions.
-Take as example a modern smartphone, some might support fingerprint authentication, some might support facial recognition authentication and some might support none of those.\\
+Take as example a modern smartphone, some might support fingerprint authentication, some might support facial recognition authentication and some support neither.\\
 	\item \textit{Variation in data.} Products might vary in the definitions of data structures.
-For example a software managing planned repairs of industrial machines might need store additional information for some companies like an SAP Id in case these planned repairs are also concurrently managed in SAP, while some other companies might not need that information.\\
+For example a software managing planned repairs of industrial machines might need store additional information for some companies like an SAP id in case these planned repairs are also concurrently managed in SAP, while some other companies might not need that information.\\
 	\item \textit{Variation in technology.} The technology different products run on may vary.
@@
 \section{Variability modeling}
-As mentioned there are two main approaches for managing variability.
-For each of which we will discuss the advantages and disadvantages and create a broad overview over the different modeling approaches.\\\\
+As mentioned, there are two main approaches for managing variability.
+For each of which we will discuss the advantages and disadvantages and create a broad overview over the different modeling approaches.\\\\\\
 \textbf{Feature modeling}\\
@@
 They can be used to define commonalities and differences between systems and their variants, as well as determine mandatory, optional or alternative aspects of them\cite{b8}.
-A broadly used approach of displaying features and their relations is hierarchically structured diagram which is called a features diagrams.\\
+A broadly used approach of displaying features and their relations is a hierarchically structured diagram which is called a features diagram.\\
 
 A big advantage of feature models, especially in the shape of feature diagrams, is a possible communication enhancement between stakeholders.
-This can be achieved by providing a overview of all functionalities which are offered to the user by the system.
-Due to the nature of the model regarding structure and due to the fact that it displays if and which planned features are already provided by another system or system family, this may improve the cost and time estimation for developing the software system.
+This can be achieved by providing an overview of all functionalities which are offered to the user by the system.
+Due to the nature of the structure of the model and due to the fact that it displays if and which planned features are already provided by another system or system family, this may improve the cost and time estimation for developing the software system.
 \\
@@
 a feature can only be choosen as part of a feature sets).
-Alternative features are can be pictured as a choice box, so one can choose \textit{one} item from a feature set.\\
+Alternative features are can be pictured as a choice box, exactly \textit{one} item must be chosen from a feature set.\\
 
@@
 The previous elaborated mandatory features are the engine and seats.
-Furthermore there is the option to install a cruise control and the option for a turbocharger these features are completely independet and optional.
+Furthermore there is the option to install a cruise control and the option for a turbocharger.
+These features are completely independet and optional.
 Last but not least one can choose between normal leather seats or alcantara ones but not both.
-Which this simplified model of a car we have all three types of features modeled.\\\\
+In this simplified model of a car we have all three types of features modeled.\\\\
 \textbf{Decision modeling}\\
-In decision modeling the variability is described with decision instead of features.
+In decision modeling the variability is described with decisions instead of features.
 A variable feature in feature modeling is equal to a decision and vice versa.
-Contrary to feature modeling decision modeling is unable to describe the commonalities of systems.\\\\
-Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependicies.
+Contrary to feature modeling, decision modeling is unable to describe the commonalities of systems.\\
+
+Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependencies.
 In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy.
-\cite{b6}\\\\
-Same as feature modeling, decision modeling has been expanded over time to enable the application coverage of DM be almost the same as FM.\cite{b6}\\
+\cite{b6}\\
+
+Same as feature modeling, decision modeling has been expanded over time to enable the application coverage of DM to be almost the same as FM.\cite{b6}\\
 
@@
 	\item \textit{Linking.} This option does not exist for any technology, but for many programming languages the linking process can be manually changed. For example by switching the DLL of a module.\\
-	\item \textit{Startup} A simple option to bind variants to the system is by configuration files, that influence the features a software system exhibits.\\
+	\item \textit{Startup} A simple option to bind variants to the system is by configuration files that influence the features a software system exhibits.\\
 	\item \textit{Adaptation to runtime environment} This typically refers to plugins of a software, which can add variation to a system and can either be provided by the original developer or even third party vendors.
@@
 \section{Discussion/Challenges}
-The research field of variability had a increasing trend of published paper per year from 2000 until 2008.
-After that it stabilized between 25-30 new research paper per year from 2008 until 2011.
-One explanation of this rise in interest in the scientific community could be that the usage of variability was becoming more broadly in dynamic systems.
+The research field of variability had an increasing trend of published paper per year from 2000 until 2008.
+After that it stabilized between 25-30 new research papers per year from 2008 until 2011.
+One explanation of this rise in interest in the scientific community could be that the usage of variability was becoming more broadly used in dynamic systems.
 \cite{b4}\\
@@
 Another aspect is that feature modeling is in its core a simple concept.
-Even though its impact in the industry is pretty big considering it is the most popular concept for variability modeling in software system until today.\cite{b9} One can say that because it is such a simple concept and due to the big amount of research in this field we understand the subject pretty well by now.\\
+Even though its impact on the industry is enormous considering it is the most popular concept for variability modeling in software system as of now.\cite{b9} Due to the big amount of research having been done in this field we understand the subject reasonably well by now.\\
 
@@
 \item \textit{Lack of support for security needs} can be a factor whichs limits the use of feature modeling.\\
-\item \textit{Poor user-friendliness} can be introduced by poor handling of variability, which decreases the usebility of a system. As an example if you \\
+\item \textit{Poor user-friendliness} can be introduced by poor handling of variability, which decreases the usability of a system.\\
 \end{itemize}
@@
 There are ways to decrease the negative impact of these aspects e.g.
-correctness and consistency can be checked by various methods such as modeling guidelines customized for specific needs, together with a reviewing process where changes to the model are checked by another person to ensure the quality of the metioned parameters.\\
+correctness and consistency can be checked by various methods such as modeling guidelines customized for specific needs, together with a reviewing process where changes to the model are checked by another person to ensure the quality of the mentioned parameters.\\
 
@@
 \section{Conclussion}
-In this document we have given a short introduction to the origin of variability, how to model it and what patterns exist as well as look over related work to these topics. \\
+In this document we have given a short introduction to the origin of variability, how to model it and what patterns exist as well as look over work related to these topics. \\
 
-This information is vital to being able to manage variability, it does however, only glance over the different subjects and only provides a theoretical basis.
+This information is vital for being able to manage variability, it does however, only glance over the different subjects and merely provides a theoretical basis.
 Practical use also depends on supporting software, which widely exists especially for the task of variability modelling, as well as practical knowledge in the field, which also aids in gaining an in depth understanding of managing variability.\\

### 2021-05-02 Paper.tex
decision modeling

@@
 Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependencies.
-In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy.
-\cite{b6}\\
+In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy.\cite{b6}\\
 
 Same as feature modeling, decision modeling has been expanded over time to enable the application coverage of DM to be almost the same as FM.\cite{b6}\\
+
+As mentioned in \cite{b11} the  different approaches for decision modeling have some common core model structure.
+There is an entity which is called decision, which represents a set of all possible choices which are available at a certain point in time.\\
+
+These decision have attributes which can vary from one modeling approach to another.
+But commonly at least a decision name or identifier are used, while the other atrributes are describing the decision in such a way that the user can understand it.
+Furthermore the decision has a data type which defines the set of valid choices for decision.\cite{b11}\\
+
+With all of the different approaches it is possible to create dependencies between decisions.
+There are value contraints for defining dependencies between values, which are supported by most approaches.
+But here it has to be mentioned that all the approaches use different concepts for value contraints and generally speaking for the management of the dependencies.\cite{b11}\\
+
+Some of them are using contraints, similar to the approach used in feature modeling.
+Another approach for modeling dependencies is with rules or decisions, which are are checked when a decision is taken.\cite{b11} \\
+
+The second approach could be more intuitive for most people with a software engineering background because of the similarity how a computer program with conditional statements works.
 
@@
 For further information we refer the reader to the references, especially \cite{b2}, which gives a more in depth overview, and \cite{b6}, which compares different approaches to variability modelling.
-\newpage
 
@@
 Bosch and M.
-Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture", Amsterdam, Netherlands, 2001, pp.
-45-54
+Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture", Amsterdam, Netherlands, 2001
 \bibitem{b3} F. Bachmann, L. Bass, "Managing Variability in Software Architectures", Carnegie Mellon University (2001)
@@
 Martinez, "The state of adoption and the challenges of systematic variability management in industry", Empirical Software Engineering, vol.
-25, pp.
-1755-1797, 2020
+25, 2020
 \bibitem{b6}
@@
 Wasowski, “Cool features and tough decisions: a comparison of
-variability modeling approaches,” Proc. of the \textit{Sixth international
+variability modeling approaches”, Proc. of the \textit{Sixth international
 workshop on variability modeling of software-intensive systems
-(VaMoS '12)}, ACM, Jan 2012, pp. 173-182.
+(VaMoS '12)}, ACM, Jan 2012
 \bibitem{b7}
@@
 Philippow, "Extending Feature Diagrams with UML Multiplicities", \textit{6th World Conference on Integrated Design \& Process Technology (IDPT2002)}, vol.
-23, 2002, pp.
-1-7.
+23, 2002
+\bibitem{b11} K. Schmid, R. Rabiser, P. Grünbacher, "A comparison of decision modeling approaches in product lines", Proc. of the \textit{Fifth international
+workshop on variability modeling of software-intensive systems
+(VaMoS '11)}, ACM, Jan 2011
 \end{thebibliography}

### 2021-05-27 Paper.tex
add corrections

@@
 \begin{abstract}
-With the growing complexity of software systems, variability of said has become an increasing problem and managing it is thus a crucial part of the present-day software engineering process.
+With the growing complexity of software systems, variability of said systems has become an increasing problem and managing it is thus a crucial part of the present-day software engineering process.
 It is particularly vital if either many product versions make use of an existing architecture over a long period of time or if a variety of software systems share a common code base.
@@
 We will try to provide a basic introduction of how variability is managed in software systems by collecting information from other studies which describe the key aspects of this vast field of research.
-In the first section we are providing some base knowledge acquired previous studies.
+In the first section we are providing some base knowledge acquired from previous studies.
 After that we will lay out how one can identify the various sources of variability in software systems.
@@
 \end{enumerate}
-These, at least, are the four most common origins of variation. Of course some more do exist, but these are beyond the scope of this document.\\
+
 \section{Variability modeling}
@@
 If a feature is mandatory it has to be selected if its parent feature is selected, the type of the parent feature does not matter.
-A optional feature can be selected freely but there may be constraints from composition rules (e.g.
+An optional feature can be selected freely but there may be constraints from composition rules (e.g.
 a feature can only be choosen as part of a feature sets).
@@
 
-Some of them are using contraints, similar to the approach used in feature modeling.
+Some of them use contraints, similar to the approach used in feature modeling.
 Another approach for modeling dependencies is with rules or decisions, which are are checked when a decision is taken.\cite{b11} \\
@@
 	\item \textit{Variant Entity.} There exist several options but only one may be bound to at any given time.\\
-	\item \textit{Optional Entity.} As the name suggests, an optional entity may or may not be included in the system.\\
+	\item \textit{Optional Entity.} An optional entity may or may not be included in the system.\\
 	\item \textit{Multiple Coexisting Entities.} Similar to the variant entity but with the exception that multiple variants can be used.\\
@@
 Another aspect is that feature modeling is in its core a simple concept.
-Even though its impact on the industry is enormous considering it is the most popular concept for variability modeling in software system as of now.\cite{b9} Due to the big amount of research having been done in this field we understand the subject reasonably well by now.\\
+Even though its impact on the industry is enormous considering it is the most popular concept for variability modeling in software system as of now.\cite{b9} \\
 
@@
 \item \textit{Runtime upgradability} and the process of switching features is sometimes necessary.\\
-\item \textit{Lack of support for security needs} can be a factor whichs limits the use of feature modeling.\\
+\item \textit{Lack of support for security needs} can be a factor which limits the use of feature modeling.\\
 \item \textit{Poor user-friendliness} can be introduced by poor handling of variability, which decreases the usability of a system.\\
@@
 
-It should also be noted most of the studies in the area of FM are including some sort of tool to support the process. Only a few don't mention any form of software support.\\
+It should also be noted most of the studies in the area of FM include some sort of tool to support the process. Only a few don't mention any form of software support.\\
 
-\section{Conclussion}
+\section{Conclusion}
 In this document we have given a short introduction to the origin of variability, how to model it and what patterns exist as well as look over work related to these topics. \\

### 2021-05-28 Paper.tex
extended conclusion

@@
 \section{Conclusion}
-In this document we have given a short introduction to the origin of variability, how to model it and what patterns exist as well as look over work related to these topics. \\
+In this document we have given a short introduction to managing variability in software systems.
+We do this by establishing the origin of variability, how to model it with either feature modeling or decision modeling as well as what patterns of variability exist in addition to looking over work related to these topics and finally presenting the challenges.
+\\
+
+Variability is intruduced early in the developement of a new product and can have four main origins: variation in function, variation in data, variation in technology and variation in environment.
+This variability can then be modelled via either one of the main approaches, namely feature modeling or decision modeling.
+The modelled variability mainly fall in three categories and can vary its binding time as described in section 5.\\
 
@@
 
-For further information we refer the reader to the references, especially \cite{b2}, which gives a more in depth overview, and \cite{b6}, which compares different approaches to variability modelling.
+By now the scientific community has a solid understanding of variability in software systems and variability modelling has become a broadly used tool.
 

### 2021-06-04 Paper.tex
finish up

--- 2021-05-27 Paper.tex
@@
 There are several types of features mentioned in \cite{b8}, named \textit{mandatory, optional and alternative features}.
-Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.\cite{b9}\\
+Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.
+Additionally so called cross tree constraints can be defined to model, for example mutual exclusive features across the whole tree structure.\cite{b9}\\
 
@@
 
+\begin{figure}[h!]
+	\includegraphics[width=\columnwidth]{decision_model}
+	\caption{Simplified decision model of a car}
+\end{figure}
+Fig.
+2 uses the same model as Fig.
+1 but this time with the decision modeling approach was used.
+As described previously the decision model does not contain information about the commonalities in the system.\\
+
 \section{Variability patterns}
@@
 These mechanisms strive to bind a given system to one out of many alternatives at different binding times.
-Binding times describe when exactly these mechanisms stop deffering the decision and make the choice for any of the variations, as illustrated in Figure 2.\cite{b2}
+Binding times describe when exactly these mechanisms stop deffering the decision and make the choice for any of the variations, as illustrated in Figure 3.\cite{b2}
 
@@
 \begin{itemize}
-	\item \textit{Variant Entity.} There exist several options but only one may be bound to at any given time.\\
-	\item \textit{Optional Entity.} An optional entity may or may not be included in the system.\\
+	\item \textit{Variant Entity.} There exist several options but only one may be bound to at any given time.
+	\item \textit{Optional Entity.} An optional entity may or may not be included in the system.
 	\item \textit{Multiple Coexisting Entities.} Similar to the variant entity but with the exception that multiple variants can be used.\\

### 2021-06-04 Paper.tex
remove unused references

@@
 Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependencies.
-In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy.\cite{b6}\\
-
+In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy.
 Same as feature modeling, decision modeling has been expanded over time to enable the application coverage of DM to be almost the same as FM.\cite{b6}\\
@@
 \begin{itemize}
-\item \textit{Guaranteed correctness and consistency}. Authors have to urge for verification and consistency checking.\\
-\item \textit{Performance overheads} are difficult to model.\\
-\item \textit{Runtime upgradability} and the process of switching features is sometimes necessary.\\
-\item \textit{Lack of support for security needs} can be a factor which limits the use of feature modeling.\\
-\item \textit{Poor user-friendliness} can be introduced by poor handling of variability, which decreases the usability of a system.\\
+\item \textit{Guaranteed correctness and consistency}. Authors have to urge for verification and consistency checking.
+\item \textit{Performance overheads} are difficult to model.
+\item \textit{Runtime upgradability} and the process of switching features is sometimes necessary.
+\item \textit{Lack of support for security needs} can be a factor which limits the use of feature modeling.
+\item \textit{Poor user-friendliness} can be introduced by poor handling of variability, which decreases the usability of a system.
 \end{itemize}
@@
 \begin{thebibliography}{00}
-\bibitem{b1} S. Apel, D. Batory, C. Kästner and G. Saake, "Software Product Lines. In: Feature-Oriented Software Product Lines", Berlin Heidelberg, Springer (2013)
 \bibitem{b2} J.
@@
 (VaMoS '12)}, ACM, Jan 2012
-\bibitem{b7}
-Software Productivity Consortium Services Corporation,
-Technical Report SPC-92019-CMC. \textit{Reuse-Driven Software
-Processes Guidebook}, Version 02.00.03, 1993.
 \bibitem{b8} K.
